PLAE web app enables powerful searching and multiple visualizations across one million unified single-cell ocular transcriptomes

### 2022-09-09 manuscript_00.Rmd
citation fix

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(glue)
library(tictoc)
#tic()()
#data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
#data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
data_dir_vGiga <- '~/data/scEiaD_gigascience/'
data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

@@

#data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
#data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
data_dir_vGiga <- '~/data/scEiaD_gigascience/'
data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vGiga}/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite'))
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vPLAE}/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20__pointRelease01.sqlite'))

@@

The retina is composed of six major cell populations: the photoreceptors, horizontal, bipolar, amacrine, ganglion, and non-neuronal cells [@masland_cell_2011]. These six populations can be further divided into dozens of cell types [@yan_cell_2020]. Furthermore, human disease of the eye can derive from damage to specific cell types. Glaucoma is characterized by damage to the retinal ganglion cells while cone and cone-rod dystrophies are characterized by photoreceptor degeneration. Efforts to untangle the transcriptomics of these diverse cell types with single cell transcriptomics date to the early 2000s when researchers [@hagstromConePigmentGene2000b; @jabsEvidenceP2XP2X2000; @liLocalizationDopamineD1receptor2004; @lindqvistSingleCellRTPCR2002a; @reidRetinoschisinPhotoreceptorsecretedProtein2003]  used RT-PCR approaches to amplify select transcripts in small numbers of cells. Later, researchers used microarray platforms to analyze a broader array of transcripts[@cherryDevelopmentDiversificationRetinal2009; @gustincichGeneDiscoveryGenetically2004a; @kimIdentificationMolecularMarkers2008a; @roeschTranscriptomeRetinalMuller2008; @trimarchiIndividualRetinalProgenitor2008; @trimarchiMolecularHeterogeneityDeveloping2007b; @wahlinMethodAnalysisGene2004a]

>>>>

```{r Setup..., message=FALSE, warning=FALSE, include=FALSE}
#knitr::opts_chunk$set(fig.pos = 'p') # Places figures on their own pages
knitr::opts_chunk$set(out.width = '100%', dpi=300)
library(tidyverse)
library(citr)
library(cowplot)
library(ggrepel)
library(colorspace)
library(flextable)
library(captioner)
library(glue)
library(tictoc)
#tic()()
#data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
#data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
#data_dir_vGiga <- '~/data/scEiaD_gigascience/'
#data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
data_dir_vGiga <- '/Volumes/Thunder/data/scEiaD/'
data_dir_vPLAE <- '/Volumes/Thunder/data/scEiaD_2022_02/'
# setup caption-ing
fig_cap <- captioner("Figure")
supFig_cap <- captioner("Supplemental Figure")
tab_cap <- captioner("Table")
supTab_cap <- captioner("Supplemental Table")

@@

#data_dir_vGiga <- '/Volumes/McGaughey_S/data/scEiaD/'
#data_dir_vPLAE <- '/Volumes/McGaughey_S/data/scEiaD_2022_02/'
#data_dir_vGiga <- '~/data/scEiaD_gigascience/'
#data_dir_vPLAE <- '~/data/scEiaD_2022_02/'
data_dir_vGiga <- '/Volumes/Thunder/data/scEiaD/'
data_dir_vPLAE <- '/Volumes/Thunder/data/scEiaD_2022_02/'

scEiaD_2020_v00 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vGiga}/MOARTABLES__anthology_limmaFALSE___5000-transform-counts-universe-batch-scVIprojectionSO-8-0.1-50-5.sqlite'))
scEiaD_2020_v01 <- pool::dbPool(RSQLite::SQLite(), dbname =  glue('{data_dir_vPLAE}/MOARTABLES__anthology_limmaFALSE___4000-counts-universe-study_accession-scANVIprojection-15-5-0.1-50-20__pointRelease01.sqlite'))

@@

The retina is composed of six major cell populations: the photoreceptors, horizontal, bipolar, amacrine, ganglion, and non-neuronal cells [@maslandCellPopulationsRetina2011]. These six populations can be further divided into dozens of cell types [@yanCellAtlasHuman2020a]. Furthermore, human disease of the eye can derive from damage to specific cell types. Glaucoma is characterized by damage to the retinal ganglion cells while cone and cone-rod dystrophies are characterized by photoreceptor degeneration. Efforts to untangle the transcriptomics of these diverse cell types with single cell transcriptomics date to the early 2000s when researchers [@hagstromConePigmentGene2000b; @jabsEvidenceP2XP2X2000; @liLocalizationDopamineD1receptor2004; @lindqvistSingleCellRTPCR2002a; @reidRetinoschisinPhotoreceptorsecretedProtein2003]  used RT-PCR approaches to amplify select transcripts in small numbers of cells. Later, researchers used microarray platforms to analyze a broader array of transcripts[@cherryDevelopmentDiversificationRetinal2009; @gustincichGeneDiscoveryGenetically2004a; @kimIdentificationMolecularMarkers2008a; @roeschTranscriptomeRetinalMuller2008; @trimarchiIndividualRetinalProgenitor2008; @trimarchiMolecularHeterogeneityDeveloping2007b; @wahlinMethodAnalysisGene2004a]
